\textbf{Microbiota predict \emph{Clostridium difficile} severity in germ-free mice colonized with human feces}

### 2022-06-15 submission/manuscript.tex
add Table S1 legend to main text
update manuscript with reviewer feedback

Next, we tested this set of mice with their human-derived gut microbial
communities for susceptibility to \emph{C. difficile} infection. A
typical mouse model of CDI requires pre-treatment of conventional mice
with antibiotics, such as clindamycin, to become susceptible to \emph{C.
difficile} colonization (20, 21). However, we wanted to avoid modifying
the gut communities with an antibiotic to maintain their unique
microbial compositions and ecological relationships. Since some of these
communities came from people at increased risk of CDI, such as recent
hospitalization or antibiotic use (18), we tested whether \emph{C.
difficile} was able to infect these mice without an antibiotic
perturbation. We hypothesized that \emph{C. difficile} would be able to
colonize the mice who received their gut communities from a donor with a
perturbed community. Mice were challenged with 10\(^{3}\) \emph{C.
difficile} RT027 clinical isolate spores. The mice were followed for 10
days post-challenge, and their stool was collected and plated for
\emph{C. difficile} colony forming units (CFU) to determine the extent
of the infection. Surprisingly, communities from all donors were able to
be colonized (Figure 2). Two mice were able to resist \emph{C.
difficile} colonization, both received their community donor N1, which
may be attributed to experimental variation since this group also had
more mice. By colonizing germ-free mice with different human fecal
communities, we were able to generate diverse gut communities in mice,
which were susceptible to \emph{C. difficile} infection without further
modification of the gut community.

@@

\textbf{Microbial community members explain variation in CDI severity.}
We next interrogated the bacterial communities at the time of \emph{C.
difficile} challenge (day 0) for their relationship to infection
outcomes using linear discriminant analysis (LDA) effect size (LEfSe)
analysis to identify individual bacterial populations that could explain
the variation in disease severity. We split the mice into groups by
severity level based on moribundity or 10 days post infection (dpi)
histopathologic score for non-moribund. This analysis revealed bacterial
operational taxonomic units (OTUs) that were significantly different at
the time of challenge by the disease severity (Figure 4A). OTUs
associated with \emph{Akkermansia}, \emph{Bacteroides},
\emph{Clostridium sensu stricto}, and \emph{Turicibacter} were detected
at higher relative abundances in the mice that became moribund. OTUs
associated with \emph{Anaerotignum}, \emph{Enterocloster}, and
\emph{Murimonas} were more abundant in the non-moribund mice that would
develop only low intestinal injury. To understand the role of toxin
activity in disease severity, we applied LEfSe to identify the OTUs at
the time of challenge that most likely explain the differences between
communities that had toxin activity detected at anytime point to those
that did not (Figure 4B). An OTU associated with \emph{Bacteroides}, OTU
7, associated with the presence of toxin also associated with
moribundity. Likewise, OTUs associated with \emph{Enterocloster} and
\emph{Murimonas} that were associated with no detected toxin also
exhibited greater relative abundance in communities from non-moribund
mice with a low histopathologic score. Lastly, we tested for
correlations between the endpoint (10 dpi) relative abundances of OTUs
and the histopathologic summary score (Figure 4C). The endpoint relative
abundance of \emph{Bacteroides}, OTU 17, was positively correlated with
histopathologic score, as its day 0 relative abundance did with disease
severity (Figure 4A). A population of \emph{Bacteroides}, OTU 17, was
positively correlated with the histopathologic score and were increased
in the group of mice with detectable toxin. We also tested for
correlations between the endpoint relative abundances of OTUs and toxin
activity but none were significant. This analysis identified bacterial
populations that were associated with the variation in moribundity,
histopathologic score, and toxin.

We next determined whether, collectively, bacterial community membership
and relative abundance could be predictive of the CDI disease outcome.
We trained logistic regression models with bacterial community relative
abundance data from the day of colonization at each taxonomic rank to
predict toxin, moribundity, and histopathologic summary score. For
predicting if detectable toxin would be produced, microbial populations
aggregated by genus rank classification performed similarly as models
using lower taxonomic ranks (mean AUROC = 0.787, Figure S3). \emph{C.
difficile} increased odds of producing detectable toxin when the
community infected had less abundant populations of \emph{Monoglobus},
\emph{Akkermansia}, \emph{Extibacter}, \emph{Intestinimonas} and
\emph{Holdemania} and had more abundant populations of
\emph{Lachnospiraceae} (Figure 5A). Next, we assessed the ability of the
community to predict moribundity. Bacteria grouped by order rank
classification was sufficient to predict which mice would succumb to the
infection before the end of the experiment (mean AUROC = 0.9205, Figure
S3). Many populations contributed to an increase odds of moribundity
(Figure 5B). Populations related to \emph{Bifidobacteriales} and
\emph{Clostridia} decreased the odds of a moribund outcome. Lastly, the
relative abundances of OTUs were able to predict a high or low
histopathologic score 10 dpi (histopathologic scores were dichotomized
as in previous analysis, mean AUROC = 0.99, Figure S3). The model
identified some similar OTUs as the LEfSe analysis, such as
\emph{Murimonas} (OTU 48), \emph{Bacteroides} (OTU 7), and
\emph{Hungatella} (OTU 24). These models have shown that the relative
abundance of bacterial populations and their relationship to each other
could be used to predict the variation in moribundity, histopathologic
score, and detectable toxin of CDI.

@@

The groups of bacteria that were associated with either a higher
histopathologic score or moribundity are members of the indigenous gut
community that also have been associated with disease, often referred to
as opportunistic pathogens. Some of the populations of
\emph{Bacteroides}, \emph{Enterococcus}, and \emph{Klebsiella} that
associated with worse outcomes, have been shown to have pathogenic
potential, expand after antibiotic use, and are commonly detected in CDI
cases (66--69). In addition to these populations, \emph{Eggerthella},
\emph{Prevotellaceae} and \emph{Helicobacter}, which associated with
worse outcomes, have also been associated with intestinal inflammation
(70--72). Recently, \emph{Helicobacter hepaticus} was shown to be
sufficient to cause susceptibility to CDI in IL-10 deficient C57BL/6
mice (73). In our experiments, when \emph{Helicobacter} was present, the
infection resulted in a high histopathologic score (Figure 4C). While we
did not use IL-10 deficient mice, it is possible the bacterial community
or host response are similarly modified by \emph{Helicobacter}, allowing
\emph{C. difficile} infection and host damage. These bacteria groups
increased in severe outcomes maintained their differences throughout the
length of the experiment (Figure S4). These results agreed Aside from
\emph{Helicobacter}, these groups of bacteria that associated with more
severe outcomes did not have a conserved association between their
relative abundance and the disease severity across all mice.

Since we observed groups of bacteria that were associated with less
severe disease it may be appropriate to apply the damage-response
framework for microbial pathogenesis to CDI (74, 75). This framework
posits that disease is not driven by a single entity, rather it is an
emergent property of the responses of the host immune system, infecting
microbe, \emph{C. difficile}, and the indigenous microbes at the site of
infection. In the first set of experiments, we used the same host
background, C57BL/6 mice, the same infecting microbe, \emph{C.
difficile} RT027 clinical isolate 431, with different gut bacterial
communities. The bacterial groups in those communities were often
present in both moribund and non-moribund and across the range of
histopathologic scores. Thus, it was not merely the presence of the
bacteria but their activity in response to the other microbes and host
which affect the extent of the host damage. Additionally, while each
mouse and \emph{C. difficile} population had the same genetic
background, they too were reacting to the specific microbial community.
Different gut microbial communities can also have different effects on
the host immune responses (76). Disease severity is driven by the
cumulative effect of the host immune response and the activity of
\emph{C. difficile} and the gut bacteria. \emph{C. difficile} drives
host damage through the production of toxin. The gut microbiota can
modulate host damage through the balance of metabolic and competitive
interactions with \emph{C. difficile}, such as bacteriocin production or
mucin degradation, and interactions with the host, such as host mucus
glycosylation or intestinal IL-33 expression (15, 77). For example, low
levels of mucin degradation can provide nutrients to other community
members producing a diverse non-damaging community (78). However, if
mucin degradation becomes too great it reduces the protective function
of the mucin layer and exposes the epithelial cells. This
over-harvesting can contribute to the host damage due to other members
producing toxin. Thus, the resultant intestinal damage is the balance of
all activities in the gut environment. Host damage is the emergent
property of numerous damage-response curves, such as one for host immune
response, one for \emph{C. difficile} activity and another for
microbiome community activity, each of which are a composite curve of
the individual activities from each group, such as antibody production,
neutrophil infiltration, toxin production, sporulation, fiber and mucin
degradation. Therefore, while we have identified populations of
interest, it may be necessary to target multiple types of bacteria to
reduce the community interactions contributing to host damage.

@@

\textbf{Statistical analysis and modeling.} To compare community
structures, we calculated Yue and Clayton dissimilarity matrices
(\(\theta\)\textsubscript{YC}) in mothur (84). For this calculation, we
averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence
reads per sample to limit uneven sampling biases. We tested for
differences in individual taxonomic groups that would explain the
outcome differences with LEfSe (85) in mothur (default parameters, LDA
\textgreater{} 4). We tested for differences in temporal trends through
fitting a linear model to each OTU and testing for differences between
histopathological summary scores with LEfSe (85) in mothur (default
parameters, LDA \textgreater{} 3). Remaining statistical analysis and
data visualization was performed in R (v4.0.5) with the tidyverse
package (v1.3.1). We tested for significant differences in
\(\beta\)-diversity (\(\theta\)\textsubscript{YC}), histopathological
scores, and toxin activity using the Wilcoxon rank sum test,
non-unimodality to non-moribund histopathological summary score using
Hartigans' dip test, and toxin detection in mice using the Pearson's
Chi-square test. We used Spearman's correlation to identify which OTUs
that had a correlation between their relative abundance and the
histopathologic summary score. \emph{P} values were then corrected for
multiple comparisons with a Benjamini and Hochberg adjustment for a type
I error rate of 0.05 (86). We built L2 logistic regression models using
the mikropml package (87). Sequence counts were summed by taxonomic
ranks from day 0 samples, normalized by centering to the feature mean
and scaling by the standard deviation, and features positively or
negatively correlated were collapsed into a single feature. For each L2
logistic regression model, we ran 100 random iterations using values of
1e-0, 1e1, 1e2, 2e2, 3e2, 4e2, 5e2, 6e2, 7e2, 8e2, 9e2, 1e3, 1e4 for the
L2 regularization penalty with a split of 80\% of the data for training
and 20\% of the data for testing. Lastly, we did not compare murine
communities to donor community or clinical data because germ-free mice
colonized with non-murine fecal communities have been shown to more
closely resemble the murine communities than the donor species community
(88). Furthermore, it is not our intention to make any inferences
regarding human associated bacteria and their relationship with human
CDI outcome.

\textbf{Code availability.} Scripts necessary to reproduce our analysis
and this paper are available in an online repository
(\url{https://github.com/SchlossLab/Lesniak_Severity_XXXX_2022}).

@@

\textbf{Figure 5. Fecal bacterial community members of the murine gut at
the time of \emph{C. difficile} infection predicted outcomes of the
infection.} On the day of infection (Day 0), bacterial community members
grouped by different classification rank were modeled with logistic
regression to predict the infection outcome. The models used the highest
taxonomic classification rank without a decrease in performance. Models
used all community members but plotted are those members with a mean
odds ratio not equal to 1. Median (solid points) and interquartile range
(lines) of the group relative abundance are plotted. Bacterial groups
are ordered by their odds ratio. * indicates that the bacterial group
was unclassified at lower taxonomic classification ranks. (A) Bacterial
members grouped by genus predicted which mice would have toxin activity
detected at any point throughout the infection (Toxin +, dark purple).
(B) Bacterial members grouped by order predicted which mice would become
moribund (dark blue). (C) Bacterial members grouped by OTU predicted if
the mice would have a high (score greater than the median score of 5,
colored dark green) or low (score less than the median score of 5,
colored light green) histopathologic summary score.

@@

\textbf{Figure S4. Temporal dynamics of OTUs that differed between
histopathologic summary score.} Relative abundance of OTUs on each day
relative to the time of \emph{C. difficile} challenge (Day 0) that have
a significantly different temporal trend by the histopathologic summary
score by LEfSe analysis. Median (points) and interquartile range (lines)
are plotted. Points and lines are colored by infection outcome of
moribund (colored black) or non-moribund with either a high
histopathologic score (score greater than the median score of 5, colored
green) or a low histopathologic summary score (score less than the
median score of 5, colored light green).

>>>>

Next, we tested this set of mice with their human-derived gut microbial
communities for susceptibility to \emph{C. difficile} infection. A
typical mouse model of CDI requires pre-treatment of conventional mice
with antibiotics, such as clindamycin, to become susceptible to \emph{C.
difficile} colonization (20, 21). However, we wanted to avoid modifying
the gut communities with an antibiotic to maintain their unique
microbial compositions and ecological relationships. Since some of these
communities came from people at increased risk of CDI, such as recent
hospitalization or antibiotic use (18), we tested whether \emph{C.
difficile} was able to infect these mice without an antibiotic
perturbation. We hypothesized that \emph{C. difficile} would be able to
colonize the mice who received their gut communities from a donor with a
perturbed community. Mice were challenged with 10\(^{3}\) \emph{C.
difficile} RT027 clinical isolate spores. The mice were followed for 10
days post-challenge, and their stool was collected and plated for
\emph{C. difficile} colony forming units (CFU) to determine the extent
of the infection. Surprisingly, communities from all donors were able to
be colonized (Figure 2). Two mice were able to resist \emph{C.
difficile} colonization, both received their community from Donor N1,
which may be attributed to experimental variation since this group also
had more mice. By colonizing germ-free mice with different human fecal
communities, we were able to generate diverse gut communities in mice,
which were susceptible to \emph{C. difficile} infection without further
modification of the gut community.

@@

\textbf{Microbial community members explain variation in CDI severity.}
We next interrogated the bacterial communities at the time of \emph{C.
difficile} challenge (day 0) for their relationship to infection
outcomes using linear discriminant analysis (LDA) effect size (LEfSe)
analysis to identify individual bacterial populations that could explain
the variation in disease severity. We split the mice into groups by
severity level based on moribundity or 10 days post infection (dpi)
histopathologic score for non-moribund mice. This analysis revealed
bacterial operational taxonomic units (OTUs) that were significantly
different at the time of challenge by the disease severity (Figure 4A).
OTUs associated with \emph{Akkermansia}, \emph{Bacteroides},
\emph{Clostridium sensu stricto}, and \emph{Turicibacter} were detected
at higher relative abundances in the mice that became moribund. OTUs
associated with \emph{Anaerotignum}, \emph{Enterocloster}, and
\emph{Murimonas} were more abundant in the non-moribund mice that would
develop low intestinal injury. To understand the role of toxin activity
in disease severity, we applied LEfSe to identify the OTUs at the time
of challenge that most likely explain the differences between
communities that had toxin activity detected at anytime point to those
that did not (Figure 4B). An OTU associated with \emph{Bacteroides}, OTU
7, associated with the presence of toxin also associated with
moribundity. Likewise, OTUs associated with \emph{Enterocloster} and
\emph{Murimonas} that were associated with no detected toxin also
exhibited greater relative abundance in communities from non-moribund
mice with a low histopathologic score. We tested for correlations
between the endpoint (10 dpi) relative abundances of OTUs and the
histopathologic summary score (Figure 4C). The endpoint relative
abundance of \emph{Bacteroides}, OTU 17, was positively correlated with
histopathologic score, as its day 0 relative abundance did with disease
severity (Figure 4A). The population of OTU 17 was also increased in the
group of mice with detectable toxin. We also tested for correlations
between the endpoint relative abundances of OTUs and toxin activity but
none were significant. Lastly, we tested for associations between
temporal changes and disease severity (Figure S4). Most groups of
bacteria maintained higher relative abundance, relative to the other
other outcome groups, from day 0 through the end of the experiment. This
analysis identified bacterial populations that were associated with the
variation in moribundity, histopathologic score, and toxin.

We next determined whether, collectively, bacterial community membership
and relative abundance could be predictive of the CDI disease outcome.
We trained logistic regression models with bacterial community relative
abundance data from the day of colonization at each taxonomic rank to
predict toxin, moribundity, and histopathologic summary score. We used
the highest taxonomic classification rank which performed similar to
lower ranks, which suggested the effect is associated with general
attributes of the bacterial group as opposed to specific functions of
more refined grouping. For predicting if detectable toxin would be
produced, microbial populations aggregated by genus rank classification
performed similarly as models using lower taxonomic ranks (mean AUROC =
0.787, Figure S3). \emph{C. difficile} increased odds of producing
detectable toxin when the community infected had less abundant
populations of \emph{Monoglobus}, \emph{Akkermansia}, \emph{Extibacter},
\emph{Intestinimonas} and \emph{Holdemania} and had more abundant
populations of \emph{Lachnospiraceae} (Figure 5A). Next, we assessed the
ability of the community to predict moribundity. Bacteria grouped by
order rank classification was sufficient to predict which mice would
succumb to the infection before the end of the experiment (mean AUROC =
0.9205, Figure S3). Many populations contributed to an increase odds of
moribundity (Figure 5B). Populations related to \emph{Bifidobacteriales}
and \emph{Clostridia} decreased the odds of a moribund outcome. Lastly,
the relative abundances of OTUs were able to predict a high or low
histopathologic score 10 dpi (histopathologic scores were dichotomized
as in previous analysis, mean AUROC = 0.99, Figure S3). The model
identified some similar OTUs as the LEfSe analysis, such as
\emph{Murimonas} (OTU 48), \emph{Bacteroides} (OTU 7), and
\emph{Hungatella} (OTU 24). These models have shown that the relative
abundance of bacterial populations and their relationship to each other
could be used to predict the variation in moribundity, histopathologic
score, and detectable toxin of CDI.

@@

The groups of bacteria that were associated with either a higher
histopathologic score or moribundity are members of the indigenous gut
community that also have been associated with disease, often referred to
as opportunistic pathogens. Some of the populations of
\emph{Bacteroides}, \emph{Enterococcus}, and \emph{Klebsiella} that
associated with worse outcomes, have been shown to have pathogenic
potential, expand after antibiotic use, and are commonly detected in CDI
cases (66--69). In addition to these populations, \emph{Eggerthella},
\emph{Prevotellaceae} and \emph{Helicobacter}, which associated with
worse outcomes, have also been associated with intestinal inflammation
(70--72). Recently, \emph{Helicobacter hepaticus} was shown to be
sufficient to cause susceptibility to CDI in IL-10 deficient C57BL/6
mice (73). In our experiments, when \emph{Helicobacter} was present, the
infection was more likely to result in a high histopathologic score
(Figure 4C, S4). While we did not use IL-10 deficient mice, it is
possible the bacterial community or host response are similarly modified
by \emph{Helicobacter}, allowing \emph{C. difficile} infection and host
damage. Aside from \emph{Helicobacter}, these groups of bacteria that
associated with more severe outcomes did not have a conserved
association between their relative abundance and the disease severity
across all mice.

Since we observed groups of bacteria that were associated with less
severe disease it may be appropriate to apply the damage-response
framework for microbial pathogenesis to CDI (74, 75). This framework
posits that disease is not driven by a single entity, rather it is an
emergent property of the responses of the host immune system, infecting
microbe, \emph{C. difficile}, and the indigenous microbes at the site of
infection. In this set of experiments, we used the same host background,
C57BL/6 mice, the same infecting microbe, \emph{C. difficile} RT027
clinical isolate 431, with different gut bacterial communities. The
bacterial groups in those communities were often present in both
moribund and non-moribund mice and across the range of histopathologic
scores. Thus, it was not merely the presence of the bacteria but their
activity in response to the other microbes and host which affect the
extent of the host damage. Additionally, while each mouse and \emph{C.
difficile} population had the same genetic background, they too were
reacting to the specific microbial community. Different gut microbial
communities can also have different effects on the host immune responses
(76). Disease severity is driven by the cumulative effect of the host
immune response and the activity of \emph{C. difficile} and the gut
bacteria. \emph{C. difficile} drives host damage through the production
of toxin. The gut microbiota can modulate host damage through the
balance of metabolic and competitive interactions with \emph{C.
difficile}, such as bacteriocin production or mucin degradation, and
interactions with the host, such as host mucus glycosylation or
intestinal IL-33 expression (15, 77). For example, low levels of mucin
degradation can provide nutrients to other community members producing a
diverse non-damaging community (78). However, if mucin degradation
becomes too great it reduces the protective function of the mucin layer
and exposes the epithelial cells. This over-harvesting can contribute to
the host damage due to other members producing toxin. Thus, the
resultant intestinal damage is the balance of all activities in the gut
environment. Host damage is the emergent property of numerous
damage-response curves, such as one for host immune response, one for
\emph{C. difficile} activity and another for microbiome community
activity, each of which are a composite curve of the individual
activities from each group, such as antibody production, neutrophil
infiltration, toxin production, sporulation, fiber and mucin
degradation. Therefore, while we have identified populations of
interest, it may be necessary to target multiple types of bacteria to
reduce the community interactions contributing to host damage.

@@

\textbf{Statistical analysis and modeling.} To compare community
structures, we calculated Yue and Clayton dissimilarity matrices
(\(\theta\)\textsubscript{YC}) in mothur (84). For this calculation, we
averaged of 1000 sub-samples, or rarified, samples to 2,107 sequence
reads per sample to limit uneven sampling biases. We tested for
differences in individual taxonomic groups that would explain the
outcome differences with LEfSe (85) in mothur (default parameters, LDA
\textgreater{} 4). We tested for differences in temporal trends through
fitting a linear regression model to each OTU and tested for differences
in regression coefficients by histopathological summary scores with
LEfSe (85) in mothur (default parameters, LDA \textgreater{} 3).
Remaining statistical analysis and data visualization was performed in R
(v4.0.5) with the tidyverse package (v1.3.1). We tested for significant
differences in \(\beta\)-diversity (\(\theta\)\textsubscript{YC}),
histopathological scores, and toxin activity using the Wilcoxon rank sum
test, non-unimodality to non-moribund histopathological summary score
using Hartigans' dip test, and toxin detection in mice using the
Pearson's Chi-square test. We used Spearman's correlation to identify
which OTUs that had a correlation between their relative abundance and
the histopathologic summary score. \emph{P} values were then corrected
for multiple comparisons with a Benjamini and Hochberg adjustment for a
type I error rate of 0.05 (86). We built L2 logistic regression models
using the mikropml package (87). Sequence counts were summed by
taxonomic ranks from day 0 samples, normalized by centering to the
feature mean and scaling by the standard deviation, and features
positively or negatively correlated were collapsed into a single
feature. For each L2 logistic regression model, we ran 100 random
iterations using values of 1e-0, 1e1, 1e2, 2e2, 3e2, 4e2, 5e2, 6e2, 7e2,
8e2, 9e2, 1e3, 1e4 for the L2 regularization penalty with a split of
80\% of the data for training and 20\% of the data for testing. Lastly,
we did not compare murine communities to donor community or clinical
data because germ-free mice colonized with non-murine fecal communities
have been shown to more closely resemble the murine communities than the
donor species community (88). Furthermore, it is not our intention to
make any inferences regarding human associated bacteria and their
relationship with human CDI outcome.

\textbf{Code availability.} Scripts necessary to reproduce our analysis
and this paper are available in an online repository
(\url{https://github.com/SchlossLab/Lesniak_Severity_mBio_2022}).

@@

\textbf{Figure 5. Fecal bacterial community members of the murine gut at
the time of \emph{C. difficile} infection predicted outcomes of the
infection.} On the day of infection (Day 0), bacterial community members
grouped by different classification rank were modeled with logistic
regression to predict the infection outcome. The models used the highest
taxonomic classification rank without a decrease in performance. Models
used all community members but plotted are those members with a mean
odds ratio not equal to 1. Median (solid points) and interquartile range
(lines) of the odds ratio are plotted. Bacterial groups are ordered by
their odds ratio. * indicates that the bacterial group was unclassified
at lower taxonomic classification ranks. (A) Bacterial members grouped
by genus predicted which mice would have toxin activity detected at any
point throughout the infection. Data with a decreased probability of
toxin activity are colored light purple and those with an increased
probability of toxin activity are colored dark purple. (B) Bacterial
members grouped by order predicted which mice would become moribund.
Data with a decreased probability of moribundity are colored light blue
and those with an increased probability of moribundity are colored dark
blue. (C) Bacterial members grouped by OTU predicted if the mice would
have a high (greater than the median score of 5) or low (less than the
median score of 5) histopathologic summary score. Data with a decreased
probability of high histopathologic score are colored light green and
those with an increased probability of high histopathologic score are
colored dark green.

@@

\textbf{Figure S4. Temporal dynamics of OTUs that differed between
histopathologic summary score.} Relative abundance of OTUs on each day
relative to the time of \emph{C. difficile} challenge (Day 0) that have
a significantly different temporal trend by the histopathologic summary
score by LEfSe analysis. Median (points) and interquartile range (lines)
of relative abundances are plotted. Points and lines are colored by
infection outcome of moribund (colored black) or non-moribund with
either a high histopathologic score (score greater than the median score
of 5, colored green) or a low histopathologic summary score (score less
than the median score of 5, colored light green).

\textbf{Table S1. Demographic information of subjects whose stool
samples used to colonize germ-free mice.}
